---
title: "Midterm Pitch"
author: "George Barzizza"
date: "2026-02-24"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
```

```{r}
url <- 'https://ourworldindata.org/grapher/government-spending-by-function.csv?v=1&csvType=full&useColumnShortNames=true'
budget <- read_csv(url)
head(budget)
url <- 'https://ourworldindata.org/grapher/fish-species-threatened.csv?v=1&csvType=full&useColumnShortNames=true'
fish <- read_csv(url)
head(fish)
url <- 'https://ourworldindata.org/grapher/threatened-endemic-mammal-species.csv?v=1&csvType=full&useColumnShortNames=true'
mammal <- read_csv(url)
head(mammal)
```

# Data Story Introduction
The *Sustainable Development Goals* (SDGs) launched in 2015 at the United Nations Sustainable Development Summit through the collaboration of all nations involved are used as an actionable framework to facilitate a more sustaible world by 2030. For my second data story, I plan to show how government spending on environmental protection differs by country relative to the amount of threatend endemic fish and mammal species by using three different datasets from [Our World in Data](https://ourworldindata.org). This data story will reveal different countries' monetary commitment to the welfare of native wildlife (i.e., fish and mammals) and, transitively, their commitment to the 14th (i.e., Life Below Water) and 15th (i.e., Life on Land) SDGs they undertook at the 2015 Summit.

### Works Cited (Datasets)
- **[budget](https://ourworldindata.org/grapher/government-spending-by-function?tab=table&tableFilter=countries)**
  
  OECD (2025) – with minor processing by Our World in Data
  
- **[fish](https://ourworldindata.org/grapher/fish-species-threatened)**
  
  FishBase (Froese and Pauly, 2008), via World Bank (2026) – processed by Our World in Data
  
- **[mammal](https://ourworldindata.org/grapher/threatened-endemic-mammal-species)**
  
  International Union for Conservation of Nature Red List of Threatened Species (2025) – with minor processing by Our World in Data
  
### Works (to be) Cited (Textual Sources)
- **[The Cost of Environmental Protection](https://www.rff.org/publications/journal-articles/the-cost-of-environmental-protection/)**
  
  (Pizer, Morgenstern & Shih 2001)
  
- **[Government Expenditures on Environmental Protection](https://visualizingenergy.org/government-expenditures-on-environmental-protection/)**
  
  (Cleveland 2025)
  
- **[The Trump Administration’s Cancellation of Funding for Environmental Protections Endangers Americans’ Health While Draining Their Wallets](https://www.americanprogress.org/article/the-trump-administrations-cancellation-of-funding-for-environmental-protections-endangers-americans-health-while-draining-their-wallets/)**
  
  (Kelly & Smith 2025)
  
- **[ESA at 50: The Destructive Cost of the ESA](https://www.fws.gov/testimony/esa-50-destructive-cost-esa)**

  (Williams 2023)
  
- **[Most funding for endangered species only benefits a few creatures. Thousands of others are left in limbo](https://apnews.com/article/endangered-species-spending-extinctions-plants-1ad806de0db9d09a38b7e82f6286c1b5)**

  (Brown & Flesher 2023)
  